---
title: "Week 7 Data Assignment"
author: "Vanessa Kellermann"
date: "2026-04-25"
output: 
  pdf_document:
  keep_tex: true
---
## Instructions
- Answer all questions in the sections below.
- Keep written responses concise (max 100 words where indicated).
- Ensure your code runs without errors.
- Knit this document before submission.

BOOOOO
---

## Setup

```{r}
#start off with a clear working environment
rm(list =ls())

# Load dataset (make sure the file is in your working directory)
data <- read.csv("dataset_5.csv")

# Inspect the data
head(data)
str(data)

# Convert variables to factors if needed
data$Population <- as.factor(data$Population)
data$Treatment <- as.factor(data$Treatment)
```

Part 1: Identify the Experimental Design 

#Answer in bullet points:

Independent variables:
Levels of each variable:
Dependent variable:
Number of replicates per group:
Type of design (e.g. continuous):

Part 2: Hypothesis 

(Max 100 words)

Write a clear, testable, and directional hypothesis.
Make sure you include both variables and provide a brief biological justification.

#Your answer:




```{r}
library(ggplot2)

ggplot(data, aes(x = Treatment,y = KnockdownTime, fill = Population)) +
  geom_boxplot() +
  labs(
    x = "Treatment",
    y = "Trait value",
    title = "Effect of Population and Treatment on Trait"
  ) +
  theme_minimal()

#Insert a new code chunk, copy the code from above  and try plotting the data 
#with Population on the x and fill = Treatment



```

Question: Which is the best way to plot the data for the question you are interested in, comparing populations next to each other or treatments next to each other? 
Explain why, hint think about your hypothesis

#Your answer:




```{r}
#Here are three different plot types, plot data and then decide which shows the patterns clearest 

#boxplot
ggplot(data, aes(x = Population, y = KnockdownTime, fill = Treatment)) +
  geom_boxplot() +
  labs(
    x = "Treatment",
    y = "Trait value",
    title = "Effect of Population and Treatment on Trait"
  ) +
  theme_minimal()

# The code below creates means and standard errors which you can then plot
library(dplyr)

summary_data <- data %>%
  group_by(Population, Treatment) %>%
  summarise(
    mean_KnockdownTime = mean(KnockdownTime, na.rm = TRUE),
    se_KnockdownTime = sd(KnockdownTime, na.rm = TRUE) / sqrt(n())
  )  
#Ignore warning

summary_data

#Bar graph

ggplot(summary_data, aes(x = Treatment, y = mean_KnockdownTime, fill = Population)) +
  geom_col(position = position_dodge(width = 0.8)) +
  geom_errorbar(aes(ymin = mean_KnockdownTime - se_KnockdownTime,
                    ymax = mean_KnockdownTime + se_KnockdownTime),
                position = position_dodge(width = 0.8),
                width = 0.2) +
  labs(x = "Treatment",
       y = "Mean Knockdown Time",
       title = "Mean Knockdown Time by Population and Treatment") +
  theme_minimal()

# Dot plot with means and standard errors

ggplot(summary_data, aes(x = Population, y = mean_KnockdownTime, color = Treatment)) +
  geom_point(size = 3) +  # Plot mean values
  geom_errorbar(aes(ymin = mean_KnockdownTime - se_KnockdownTime, ymax = mean_KnockdownTime + se_KnockdownTime), width = 0.2) +
    theme_minimal()

#Once you have plotted all three types run the code one final time using the plot
#type you think best describes the data


```

Part 4: 

Write a figure caption

##Your answer:



Interpretation 

(Max 100 words)

Describe the pattern shown in your figure
State whether it supports your hypothesis
Provide a brief biological explanation

#Your answer:









##Double check:
 All sections completed
 Code runs without errors
 Document successfully knitted

