---
title: "Week 7 Data Assignment"
author: "Student name and number"
date: "2026-04-25"
output:
  html_document:
    df_print: paged
  keep_tex: true
  pdf_document: null
---
## Instructions
- Answer all questions in the sections below (where it says ##Your answer:).
- Keep written responses concise (max 100 words where indicated).
- Ensure your code runs without errors.
- Knit this document as a pdf and submit it to the lms. Alternatively if having
trouble knitting insert answers into the word document provided

---

## Setup

```{r}
#start off with a clear working environment
rm(list =ls())

# Load dataset (make sure the file is in your working directory)
data <- read.csv("dataset_1.csv")

# Inspect the data
head(data)
str(data)

# Convert variables to factors if needed
data$Population <- as.factor(data$Population)
data$Treatment <- as.factor(data$Treatment)

table(data$Treatment, data$Population)
```

Part 1: Identify the Experimental Design 

#Answer in bullet points:

Independent variables:
Levels of each variable:
Dependent variable:
Number of replicates per group:
Type of design (e.g. continuous, factors):

Part 2: Hypothesis (Max 100 words)

Write a clear, testable, and directional hypothesis.
Make sure you include both variables and provide a brief biological justification.

#Your answer:



```{r}
library(ggplot2)

#Figure 1

ggplot(data, aes(x = Treatment,y = KnockdownTime, fill = Population)) +
  geom_boxplot() +
  labs(
    x = "Treatment",
    y = "Trait value",
    title = "Effect of Population and Treatment on Trait"
  ) +
  theme_minimal()

#Insert a new code chunk, copy the code from above and try plotting the data with Population on the x and fill = Treatment



```
Part 3:Data visualisation

There is more than one valid way to visualise these data.

Comparing populations on the x-axis, or
Comparing treatments on the x-axis?

Choose which way you would like to present your data for Figure 1.Justify your choice by 
referring to your hypothesis and explaining how your choice affects 
interpretation of the pattern in the data.

There is no single correct answer — your mark is based on how well your 
choice aligns with your hypothesis and interpretation.
##Your answer:




```{r}
#Figure 2

#Here are three different plot types, plot data and then decide which shows the patterns clearest 

#You will need to modify all three figures to change whether you are comparing population next to each other or treatments next to each other. You will need to modify this bit of code 
#aes(x = Population, y = KnockdownTime, fill = Treatment)) and labels labs(
#   x = "Treatment",
#  y = "Trait value",
#  title = "Effect of Population and Treatment on Trait")


#boxplot
ggplot(data, aes(x = Population, y = KnockdownTime, fill = Treatment)) +
  geom_boxplot() +
  labs(
    x = "Treatment",
    y = "Trait value",
    title = "Effect of Population and Treatment on Trait"
  ) +
  theme_minimal()

# The code below creates means and standard errors which you can then plot
library(dplyr)

summary_data <- data %>%
  group_by(Population, Treatment) %>%
  summarise(
    mean_KnockdownTime = mean(KnockdownTime, na.rm = TRUE),
    se_KnockdownTime = sd(KnockdownTime, na.rm = TRUE) / sqrt(n())
  )  
# Ignore the summarise() grouping warning — it does not affect your results

summary_data

#Bar graph

ggplot(summary_data, aes(x = Treatment, y = mean_KnockdownTime, fill = Population)) +
  geom_col(position = position_dodge(width = 0.8)) +
  geom_errorbar(aes(ymin = mean_KnockdownTime - se_KnockdownTime,
                    ymax = mean_KnockdownTime + se_KnockdownTime),
                position = position_dodge(width = 0.8),
                width = 0.2) +
  labs(x = "Treatment",
       y = "Mean Knockdown Time",
       title = "Mean Knockdown Time by Population and Treatment") +
  theme_minimal()

# Dot plot with means and standard errors

ggplot(summary_data, aes(x = Population, y = mean_KnockdownTime, color = Treatment)) +
  geom_point(size = 3) +  # Plot mean values
  geom_errorbar(aes(ymin = mean_KnockdownTime - se_KnockdownTime, ymax = mean_KnockdownTime + se_KnockdownTime), width = 0.2) +
    theme_minimal()

#All three plot types are valid ways to display these data. Choose one to communicate your results in relation to your hypothesis. Re-run code for your chosen graph

```

Part 3 continued: 

Write a figure caption (max 50 words)

##Your answer:



Interpretation of the figure (max 100 words) 
Describe the pattern shown in your figure
State whether it supports your hypothesis
Provide a brief biological explanation

#Your answer:







Once finished knit document. Before knitting go back to the start and make sure your code runs with no errors. Knitting runs all of the document code and will not work if there are any code errors. The knit option is in the middle of the task bar just above, it looks like a ball of wool with knitting needles, press the knit button.

If you really can't get knit to work you can copy your answer into the word document provided. See on the right handside word document entitled answer template.

##Double check:
 All sections completed
 Code runs without errors
 Document successfully knitted, if failing knit copy  answers into 
 word document on the lms (should also be in your project folder to the right). Fill in and upload to the assignment submission link

